Litrepl: Literate Paper Processor Promoting Transparency More Than Reproducibility

### 2024-12-31 paper/paper.tex
Update paper

To efficiently detect sections within a document, Litrepl employs simplified
versions of each document markup grammar, which it defines and parses. In this
approach, documents are abstracted as a straightforward sequence comprising
code, result, and text sections. Additionally, Litrepl identifies ignore blocks,
which serve as block comments. Any sections encapsulated within these ignore
blocks are excluded from parsing and evaluation. The structure of any document,
therefore, can be articulated using the following informal grammar:

The specific markers that denote each block type are determined by the document
format and configuration settings. For example, in a LaTeX document containing
Python code, Litrepl recognizes markers such as \verb|\begin|\verb|{python}| or
\verb|% |\verb|python| as code section begin markers, and
\verb|\end|\verb|{python}| or \verb|% |\verb|nopython| as end markers. These
latter variants are transparent to LaTeX, facilitating the inclusion of LaTeX
markup seamlessly.

The template grammars similar to the example are hardcoded for Markdown and
LaTeX document formats. In each run, Litrepl compiles them to the document
parser based on its configuration options using the Lark\cite{Lark} parsing
library.

>>>>

Litrepl abstracts documents as a straightforward sequence comprising code,
result, and text sections. Additionally, Litrepl identifies ignore blocks, which
act as comments that prevent enclosed sections from being evaluated.

Template grammars similar to the illustrative example above are encoded for
Markdown and LaTeX formats. Before each run, Litrepl calls Lark\cite{Lark} to
compile a customized parser and uses it to access the sections.
